---
title: "RWA analysis"
output:
  html_document:
    df_print: paged
---

The data was taken from the results of 2 decathlons and the performance of the participants. The data was cleaned by first cleaning the column names and then changing the row names into a column. The data in this column was changed to title case as they were names, and the columns of events were pivoted into 2 columns of events and the participants result in the event. 

```{r}
library(tidyverse)
library(dplyr)

source("data_cleaning_scripts/rwa_data_cleaning.R")
glimpse(rwa_clean)
```

  
What’s the average RWA score for each gender?
```{r}
rwa_clean %>% 
  group_by(gender) %>% 
  summarise(mean(total_rwa_score))
```
answer: 
Female - 49.6
Male - 60.4
Other - 36.4

What’s the average RWA score for left handed people vs. right handed people?
```{r}
rwa_clean %>% 
  filter(hand == "Left" | hand == "Right") %>% 
  group_by(hand) %>% 
  summarise(mean(total_rwa_score))
```
answer:
Left - 56.0
Right - 56.2

What’s the average family size for each type of childhood?
```{r}
rwa_clean %>% 
  group_by(urban) %>% 
  summarise(mean(familysize))
```
answer:
Rural - 2.8 children
Suburban - 2.6 children
Urban - 2.4 children

What’s the average time to take the test for each education level?
```{r}
rwa_clean %>%
  group_by(education) %>% 
  summarise(mean(testelapse))
```
answer:
University degree - 613 seconds
Graduate degree - 328 seconds
High school - 495 seconds
Less than high school - 386 seconds

Create a plot of results of question 4.
```{r}
test_time <- c(386, 495, 328, 613)

barplot(test_time, col = "pink", xlab = "education level", ylab = "time (s)",
        names.arg = c("Less than high school", " High school", "Graduate degree", "University degree"))

```
What’s the average RWA score for people aged:

- under 18
- 18 - 25
- 26 - 40
- 41 - 60
- over 60
 

```{r}
rwa_clean %>% 
  mutate(age_range = case_when(
    age < 18 ~ "less than 18",
    age > 17 & age < 25 ~ "18 - 25",
    age > 25 & age < 41 ~ "26 - 40",
    age > 40 & age < 61 ~ "41-60",
    age > 60 ~ "over 60"
  )) %>% 
  group_by(age_range) %>% 
  summarise(average_rwa_score = mean(total_rwa_score))
```
answer:

under 18 - 66.1
18 - 25 - 62.1
26 - 40 - 51.0
41 - 60 - 50.2
over 60 - 47.4


```{r}
average_rwa <- c(66.1, 62.1, 51.0, 50.2, 47.4)


barplot(average_rwa, col = "pink", xlab = "age range", ylab = "average rwa score",
     names.arg = c("under 18", "18 - 25", "26 - 40", "41-60", "over 60"))
```


