---
title: "RWA analysis"
output:
  html_document:
    df_print: paged
---

The data was taken from an online psychometric test on how much an individual identified with the right wing authoritarian ideology. The data was cleaned by cleaning the column names, removing unnecessary columns, filtering for valid test results, and then replacing the data values with their text equivalents to make the data easy to interpret. We assumed that anyone who had ticked questions vcl6, vcll9 and vcl2 were not answering truthfully and so were removed from the data.

```{r}
library(tidyverse)
library(dplyr)

source("data_cleaning_scripts/rwa_data_cleaning.R")
glimpse(rwa_clean)
```

  
What’s the average RWA score for each gender?
```{r}
rwa_clean %>% 
  group_by(gender) %>% 
  summarise(mean(total_rwa_score))
```
answer: 
Female - 49.6
Male - 60.4
Other - 36.4

What’s the average RWA score for left handed people vs. right handed people?
```{r}
rwa_clean %>% 
  filter(hand == "Left" | hand == "Right") %>% 
  group_by(hand) %>% 
  summarise(mean(total_rwa_score))
```
answer:
Left - 56.0
Right - 56.2

What’s the average family size for each type of childhood?
```{r}
rwa_clean %>% 
  group_by(urban) %>% 
  summarise(mean(familysize))
```
answer:
Rural - 2.8 children
Suburban - 2.6 children
Urban - 2.4 children

What’s the average time to take the test for each education level?
```{r}
rwa_clean %>%
  group_by(education) %>% 
  summarise(mean(testelapse))
```
answer:
University degree - 613 seconds
Graduate degree - 328 seconds
High school - 495 seconds
Less than high school - 386 seconds

Create a plot of results of question 4.
```{r}
test_time <- c(386, 495, 328, 613)

barplot(test_time, col = "pink", xlab = "education level", ylab = "time (s)",
        names.arg = c("Less than high school", " High school", "Graduate degree", "University degree"))

```
What’s the average RWA score for people aged:

- under 18
- 18 - 25
- 26 - 40
- 41 - 60
- over 60
 

```{r}
rwa_clean %>% 
  mutate(age_range = case_when(
    age < 18 ~ "less than 18",
    age > 17 & age < 25 ~ "18 - 25",
    age > 25 & age < 41 ~ "26 - 40",
    age > 40 & age < 61 ~ "41-60",
    age > 60 ~ "over 60"
  )) %>% 
  group_by(age_range) %>% 
  summarise(average_rwa_score = mean(total_rwa_score))
```
answer:

under 18 - 66.1
18 - 25 - 62.1
26 - 40 - 51.0
41 - 60 - 50.2
over 60 - 47.4


```{r}
average_rwa <- c(66.1, 62.1, 51.0, 50.2, 47.4)


barplot(average_rwa, col = "pink", xlab = "age range", ylab = "average rwa score",
     names.arg = c("under 18", "18 - 25", "26 - 40", "41-60", "over 60"))
```


