---
title: "Decathlon data analysis"
output: html_notebook
---

The data was taken from the results of 2 decathlons and the performance of the participants. The data was cleaned by first cleaning the column names and then
changing the row names into a column. The data in this column was changed to title case as they were names, and the columns of events were pivoted into 2 columns
of events and the participants result in the event.

```{r}
source("clean_data/Decathlon_data_clean.R")
```

```{r}
glimpse(decathlon_clean)
```

Who has the longest long jump?
```{r}
decathlon_clean %>% 
  filter(event == "long_jump") %>%
  arrange(event) %>% 
  head(1)
```
What was the avg 100m time in each competition?
```{r}
decathlon_clean %>% 
  filter(event == "100m") %>% 
  group_by(competition) %>% 
  summarise(mean_time_s = mean(result))
```
Who has the highest total points across both competitions?

```{r}
decathlon_clean %>% 
  group_by(name) %>% 
  summarise(total_points = sum(points)) %>% 
  slice_max(total_points, n = 1)
```
What was the shotput scores for the top three competitors in each competition?
```{r}
decathlon_clean %>% 
  filter(event == "shot_put") %>% 
  group_by(competition) %>% 
  slice_max(points, n = 3)
```

What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}
decathlon_clean %>% 
  filter(event == "400m") %>% 
  group_by("time_<_50s" = result < 50) %>% 
  summarise(average_points = mean(points))
```

