---
title: "Task 4 - Halloween candy analysis"
output: html_notebook
---

The data was taken from an online survey on whether participants would feel "joy","despair" or "meh" at getting certain items in their trick or treating bag at Halloween. It came in 3 different tables from 3 years (2015, 2016 and 2017). The data was cleaned by cleaning the column names of the individual tables and ensuring they match, along with removing unnecessary data. The tables were then joined and some columns were aggregated for easier reading, and checked for outlying values. Finally, more unnecessary columns were removed, and the types of candy columns were pivoted to create two columns, type of candy and rating. When aggregating the different inputs for "country" some of the inputs had to be assumed to be in reference to a certain country. It was also assumed a lot of items were joke questions and we would not actually expect anyone to give them out to trick or treaters at halloween, and were thus removed.

```{r}
library(tidyverse)
library(dplyr)

source("data_cleaning_scripts/Halloween_candy_cleaning.R")
```

What is the total number of candy ratings given across the three years?

```{r}
candy_all_clean %>% 
  drop_na(rating) %>% 
  nrow()
```
answer: 583964 ratings

What was the average age of people who are going out trick or treating?
```{r}
candy_all_clean %>% 
  drop_na(age) %>% 
  summarise(mean(age))
```
answer: 38.7 years old

What was the average age of people who are not going trick or treating?
```{r}
candy_all_clean %>% 
  group_by(going_trick_or_treating) %>% 
  summarise(mean(age))
```
answer: 39.1 years old


For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
candy_all_clean %>% 
  drop_na(rating) %>% 
  group_by(candy, rating) %>% 
  summarise(rating_count = n()) %>% 
  group_by(rating) %>% 
  slice_max(rating_count, n = 1)
```
answer: most Joys: any full sized candy bar - 7243
most mehs: lollipop - 1507
most despairs: gum from baseball card - 6989

How many people rated Starburst as despair?
```{r}
candy_all_clean %>% 
  filter(candy == "starburst", rating == "DESPAIR") %>% 
  nrow()
```
answer: 1865 people rated starburst as despair

rating system: +1 for joy, 0 for meh and -1 for despair 

What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_clean_rating <- candy_all_clean %>% 
  mutate(rating_value = case_when(
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0,
    rating == "DESPAIR" ~ -1
))
```

```{r}
candy_clean_rating %>%
  drop_na(rating) %>% 
  group_by(gender, candy) %>% 
  summarise(total_rating = sum(rating_value)) %>% 
  group_by(gender) %>% 
  slice_max(total_rating, n = 1)
```
answer:
Female - any full sized candy bar - 855
Male - any full sized candy bar - 1533
I'd rather not say - any full sized candy bar - 48
Other - any full sized candy bar - 35

What was the most popular candy bar in each year?
```{r}
candy_clean_rating %>%
  drop_na(rating) %>% 
  group_by(year, candy) %>% 
  summarise(total_rating = sum(rating_value)) %>% 
  group_by(year) %>% 
  slice_max(total_rating, n = 1)
```
answer: any full sized candy bar - all years

What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
candy_clean_rating %>%
  drop_na(rating) %>% 
  group_by(country, candy) %>% 
  summarise(total_rating = sum(rating_value)) %>% 
  group_by(country) %>% 
  slice_max(total_rating, n = 1)
```
answer: 
USA - any full sized candy bar - 2036
UK - toblerone something or other - 25
Canada - mars - 42
other - any full sized candy bar - 4768
